Offline Reinforcement Learning with Tsallis Regularization

### 2023-04-20 paper/main.tex
added main theoretical results

A more general form of entropy known as Tsallis entropy is proposed by \cite{TsallisEntropy},  defined based on the \qlog that is referred to as \emph{deformed logarithm} \cite{tsallis2009introduction}.
Let us revisit the definition of \qlog its unique inverse function $q$-exponential (analogous to the exponential function), see Figure \ref{fig:q_stats}.
We can define the Tsallis entropy $S_q(\pi)$ in a similar way to Shannon entropy with \qlog:
\begin{align}
    \begin{split}
        q \in  \mathbb{R}, \quad \logq{x} = \frac{x^{q-1} - 1}{q-1}, \quad \expq{x} = \AdaRectBracket{1 + (q - 1)x}^{\frac{1}{q-1}}_{+},\quad \tsallis{\pi} = -\AdaAngleProduct{\pi}{\logq{\pi}}.
    \end{split}
\end{align}
When $q=2$, we recover the Tsallis sparse entropy \cite{Martins16-sparsemax}, which achieves the sparsity by projecting onto the probability simplex \cite{Blondel-2020LearningFenchelYoundLoss}.
In RL, it has been investigated by \cite{Lee2018-TsallisRAL,Lee2020-generalTsallisRSS} and it is known that the optimal regularized policy takes the form:
\begin{align*}
    &\pi_{k+1}(a|s) = \AdaRectBracket{\frac{Q_k(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_k(s, \cdot)}{\tau}}}_{+} = \exp_2 \AdaBracket{\frac{Q_k(s,a)}{\tau } - \tilde{\psi}\AdaBracket{\frac{Q_k(s, \cdot)}{\tau }}}, \\
    &{\psi}\AdaBracket{\frac{Q_{k}(s,\cdot)}{\tau}} \doteq \frac{\sum_{a\in K(s)} \frac{Q(s,a)}{\tau} - 1 }{|K{(s)}|}, \,\, \tilde{\psi}\AdaBracket{\frac{Q_{k}(s,\cdot)}{\tau}} := \psi\AdaBracket{\frac{Q_{k}(s,\cdot)}{\tau}} + \frac{1}{q-1}
\end{align*}
where $[\cdot]_{+} = \max \{\cdot, 0\}$,  $K(s)$ is the set of highest-value actions satisfying $1 \!+\! i\frac{Q(s,a_{(i)})}{\tau} \!>\! \sum_{j=1}^{i}\frac{Q(s,a_{(j)})}{\tau}$., with $a_{(j)}$ denotes the $j$-th largest action.

@@

 The backward Tsallis learning consists in learning $\pi_{t, \datasetPolicy} \preceq \datasetPolicy$ given $K_\mathcal{D}$.
% However, it should be noted that the learned policy no longer requires a thresholding function $\tilde{\psi}$ since the set of allowable actions $K_\mathcal{D}$ is given, hence no action need to be explicitly truncated.
% Specifically, we propose to modify the learned policy form as 
\begin{align}
    \pi_{t+1, \datasetPolicy}(a|s) = \exp_q\AdaBracket{\frac{Q_{t, \datasetPolicy}(s,a)}{\tau} - \tilde{\psi}{\AdaBracket{\frac{Q_{t, \datasetPolicy}(s,\cdot)}{\tau}}}},  \quad \sum_{a\in K_{\mathcal{D}}(s)} \pi_{t+1, \datasetPolicy} = 1.
    \label{eq:proposal_policy}
\end{align}
When $q=2$, the policy is a variant of categorical distributions, which is less susceptible to numerical issues than exponential functions \cite{Tsai2021-selfsupervisedRelativePredictiveCoding}.

@@

\textbf{Why use Tsallis backward learning? }
If the Tsallis dataset policy assumption holds, and Tsallis backward learning is adopted, 
then the KL divergence between the learned policy and the behavior policy can be bounded as:
\begin{align}
    \begin{split}
        &\KLany{\pi_t(\cdot|s)}{\datasetPolicy(\cdot|s)} = \expectation{a\sim\pi_{t}(\cdot | s)}{\ln \pi_{t}(a|s) - \ln\datasetPolicy(a|s)} \\
        & = \expectation{a\sim\pi_{t}(\cdot | s)}{\underbrace{\ln \pi_{t}(a|s) - \ln_q\pi_{t}(a|s)}_{(1)} + \underbrace{\ln_q\pi_{t}(a|s) -  \ln\datasetPolicy(a|s)}_{(2)} + \underbrace{\ln_q\datasetPolicy(a|s) - \ln\datasetPolicy(a|s)}_{(3)} } .
    \end{split}
\end{align}
Let us now respectively bound the three terms:
\begin{align}
    \begin{split}
        (1): &\ln \pi_{t}(a|s) - \ln_q\pi_{t}(a|s) = (q-1) \AdaRectBracket{\frac{d}{dq}\ln_q \pi_t(a|s) + \ln_q\pi_t(a|s)\ln\pi_t(a|s)}\\
        &= (q-1)\AdaRectBracket{\pi_t^{q-1}(a|s)\ln\pi_t(a|s) - \frac{\pi_{t}^{q-1}(a|s)-1}{(q-1)^2} + \ln_q\pi_t(a|s)\ln\pi_t(a|s)}\\
        &= (q-1)\pi_{t}^{q-1}(a|s)\ln\pi_t(a|s) - \ln_q\pi_t(a|s) + (q-1)\ln_q\pi_t(a|s)\ln\pi_t(a|s)\\
        % &= (q-1)\pi_{t}^{q-1}(a|s)\ln\pi_t(a|s) - \AdaBracket{ \frac{Q_{t-1}(s,a)}{\tau} - \tilde{\psi}\AdaBracket{\frac{Q_{t-1}(s,\cdot)}{\tau}}}  + \ln\pi_t(a|s) \AdaBracket{ \frac{Q_{t-1}(s,a)}{\tau} - \tilde{\psi}\AdaBracket{\frac{Q_{t-1}(s,\cdot)}{\tau}}}
        & \leq (q-1)\pi_{t}^{q-1}(a|s)\ln\pi_t(a|s) + \frac{1}{q-1} - \ln\pi_t (a|s)\\
        &=  \AdaBracket{ (q-1)\pi_{t}^{q-1}(a|s) - 1}\ln\pi_t(a|s) + \frac{1}{q-1},
    \end{split}
\end{align}
where we leveraged the fact that $-\frac{1}{q-1} < \frac{Q_{t-1}(s,a)}{\tau} - \tilde{\psi}\AdaBracket{\frac{Q_{t-1}(s,\cdot)}{\tau}} < 0, \, \forall a\in K(s)$. If $a \notin K_{\mathcal{D}}(s)$, then $\ln\pi_t(a|s) = -\infty$ and the KL term is unbounded.
We repeat the same trick to yield an upper bound $\AdaBracket{ (q-1)\datasetPolicy^{q-1}(a|s) - 1}\ln\datasetPolicy(a|s) + \frac{1}{q-1}$  for (3); and and upper bound $\frac{1}{q-1}$  for (2).

As a summary, we can state for $q>1$ that:
\begin{align}
    \begin{split}
    &\KLany{\pi_t(\cdot|s)}{\datasetPolicy(\cdot|s)}  \leq  \\
    &\quad K_{\mathcal{D}}(s){\AdaRectBracket{\AdaBracket{ (q-1) \datasetPolicy^{q-1}(a|s) - 1}\ln\datasetPolicy(a|s) + \AdaBracket{  (q-1) \pi_{t}^{q-1}(a|s) - 1}\ln\pi_t(a|s) + \frac{3}{q-1}}}.
\end{split}
\end{align}
This result suggests that in order to keep the learned policy close to the behavior policy, one should choose $q$ such that $(q-1)\pi^{q-1}_t(a|s)$ close to 1.

@@

Furthermore, the assumption naturally leads to the choice of Tsallis regularized backward learning which recovers in-sample softmax as a special case when $q=1$.
Since softmax policies have full support, there always exists non-negligible probabilities for suboptimal actions; while for \eq{\ref{eq:proposal_policy}} the optimal policy can switch between greedy policy when $Q_{t, \datasetPolicy}(s,a) \leq 0, \forall a\neq a^*$ and multimodal policy when some actions have positive values.

@@

We follow the same setup here, but replacing every appearance of $\ln, \exp$ to their \qlog and $q$-exponential counterpart.
In practice, computing the thresholding function $\psi$ requires sorting in the discrete control setting and it remains unclear how to accurately estimate it for the continuous control problems.
We found it improves the performance by simply removing $\psi$ and renormalize the policy.
In this case, thresholding can still be achieved by simply learning $Q_{\theta} \rightarrow 0$.
That is, the learned policy form becomes $\pi_{t+1}(a|s) \propto \AdaRectBracket{Q_{\theta}(s,a)}_{+}$.
In order to also sample actions from the dataset when updating the policy, we repeat the derivation in \eq{\ref{eq:insample_policy}} but for the sparsemax policy.
% Let us define $q_{\theta}(s,a) := \frac{Q_{\datasetPolicy, k}(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_{\datasetPolicy, k}(s,\cdot)}{\tau}}$ for convenience:
\begin{align}
    \begin{split}
        \pi_{\datasetPolicy, k+1} (a|s) &= \datasetPolicy(a|s) \datasetPolicy(a|s)^{-1} \exp_q{Q_{\theta}(s,a)} \\
        &= \datasetPolicy(a|s)  \exp_q{\AdaBracket{\ln_2{\frac{1}{\datasetPolicy(a|s)}}}} \exp_q{Q_{\theta}(s,a)}\\
        &=  \datasetPolicy(a|s) \AdaBracket{ \exp_q\AdaBracket{Q_{\theta}(s,a) + \ln_2\frac{1}{\datasetPolicy(a|s)}} -  Q_{\theta}(s,a)  \ln_2{\frac{1}{\datasetPolicy(a|s)}} }.
        \label{eq:tsallis_inac_policy}
    \end{split}
\end{align}
In the last step we made use of the relationship $\AdaBracket{\expq{x}\cdot \expq{y}}^{q-1} = \expq{\AdaBracket{x+y}}^{q-1} + (q-1)^2 xy$ \cite{Yamano2004-properties-qlogexp}.
Similar to the discussion after \eq{\ref{eq:inac_policy}}, the Tsallis policy we derived here can be seen as the result from Tsallis KL regularization, plus another regularization that gives rise to the additional term inside the $\exp_q$ function.

% In the discrete case, the truncation threshold or the normalization function $\psi$ is computed by sorting the action values.
% However, in the continuous case we can no longer do that.
% Instead, we propose to use a single-point estimate:
% \begin{align*}
%     &V(s) \approx \frac{1}{2}\tau \AdaBracket{ \AdaBracket{\frac{Q_{\datasetPolicy, k}(s',a')}{\tau}}^2 - \psi \AdaBracket{\frac{Q_{\datasetPolicy, k}(s', \cdot)}{\tau}}^2 + 1} \\
%     &\Leftrightarrow \psi\AdaBracket{\frac{Q_{\datasetPolicy, k}(s,\cdot)}{\tau}} \approx \sqrt{\AdaBracket{\frac{Q_{\datasetPolicy, k}(s',a')}{\tau}}^2 - \frac{V(s) - \frac{1}{2}\tau}{\frac{1}{2}\tau}}.
% \end{align*}
% {\color{red}(need to record this $\psi$ value in the experiment to see how big area of the action Gaussian distribution is truncated.)}
% The single-point estimate is similar to the common practice of entropy-regularized actor-critic algorithms such as soft actor-critic \cite{haarnoja-SAC2018} that approximates the value function by a single-point estimate, i.e. $V_{\zeta}(s) \approx Q_{\theta}(s,a) + \ln\pi_{\phi}(a|s)$.

We now summarize the loss functions for implementing the Tsallis in-sample actor-critic algorithm:
\begin{align}
    % & \psi\AdaBracket{\frac{Q_{\datasetPolicy, \theta}(s,\cdot)}{\tau}} \leftarrow \sqrt{\AdaBracket{\frac{Q_{\theta}(s,a)}{\tau}}^2 - \frac{V_{\zeta}(s) - \frac{\tau}{2}}{\frac{\tau}{2}} }, \\ 
    % & Q_{\theta}(s,a) \leftarrow \frac{Q_{\theta}(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_{\theta}(s,\cdot)}{\tau}},\\
    &\mathcal{L}_\text{actor}(\phi) = -\expectation{s,a \sim \mathcal{D}}{ \AdaBracket{ \exp_q\AdaBracket{Q_{\theta}(s,a) + \ln_2\frac{1}{\pi_{\omega}(a|s)}} -  Q_{\theta}(s,a)  \logq{\frac{1}{\pi_{\omega}(a|s)}} } \ln\pi_{\phi}(a|s) }, \\
    & \mathcal{L}_\text{baseline}(\zeta) = \expectation{s\sim\mathcal{D}, a\sim \pi_{\phi}(s)}{\AdaBracket{ v_{\zeta}(s) - Q_{\theta}(s,a) - \tau\ln_2\pi_{\phi}(a|s)}^2},\\
    & \mathcal{L}_\text{critic}(\theta) = \expectation{s,a,r,s' \sim \mathcal{D}}{\AdaBracket{r + \gamma v_{\zeta}(s') - Q_{\theta}(s,a)}^2 }.
\end{align}

>>>>

A more general form of entropy known as Tsallis entropy is proposed by \cite{TsallisEntropy},  defined based on the \qlog that is referred to as \emph{deformed logarithm} \cite{tsallis2009introduction}.
Let us revisit the definition of \qlog its unique inverse function $q$-exponential (analogous to the exponential function), see Figure \ref{fig:q_stats}.
We can define the Tsallis entropy $S_q(\pi)$ in a similar way to Shannon entropy with \qlog:
\begin{align}
    \begin{split}
        q \in  \mathbb{R}, \quad \logq{x} = \frac{x^{q-1} - 1}{q-1}, \quad \expq{x} = \AdaRectBracket{1 + (q - 1)x}^{\frac{1}{q-1}}_{+},\quad \tsallis{\pi} = -\AdaAngleProduct{\pi}{\logq{\pi}}.
    \end{split}
\end{align}
When $q=2$, we recover the Tsallis sparse entropy \cite{Martins16-sparsemax}, which achieves the sparsity by projecting onto the probability simplex \cite{Blondel-2020LearningFenchelYoundLoss}.
In RL, it has been investigated by \cite{Lee2018-TsallisRAL,Lee2020-generalTsallisRSS} and it is known that the optimal regularized policy takes the form:
\begin{align*}
    &\pi_{k+1}(a|s) = \AdaRectBracket{\frac{Q_k(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_k(s, \cdot)}{\tau}}}_{+} = \exp_2 \AdaBracket{\frac{Q_k(s,a)}{\tau } - \tilde{\psi}\AdaBracket{\frac{Q_k(s, \cdot)}{\tau }}}, \\
    &{\psi}\AdaBracket{\frac{Q_{k}(s,\cdot)}{\tau}} \doteq \frac{\sum_{a\in K(s)} \frac{Q(s,a)}{\tau} - 1 }{|K{(s)}|}, \,\, \tilde{\psi}\AdaBracket{\frac{Q_{k}(s,\cdot)}{\tau}} := \psi\AdaBracket{\frac{Q_{k}(s,\cdot)}{\tau}} + \frac{1}{q-1}
\end{align*}
where $[\cdot]_{+} = \max \{\cdot, 0\}$,  $K(s)$ is the set of highest-value actions satisfying $1 \!+\! i\frac{Q(s,a_{(i)})}{\tau} \!>\! \sum_{j=1}^{i}\frac{Q(s,a_{(j)})}{\tau}$., with $a_{(j)}$ denotes the $j$-th largest action.
{\color{red}(say something here about the general $q$ cases, and how $q \rightarrow \infty$ the policy tends to uniform distribution, provide proof.)}

@@

 The backward Tsallis learning consists in learning $\pi_{t, \datasetPolicy} \preceq \datasetPolicy$ given $K_\mathcal{D}$.
% However, it should be noted that the learned policy no longer requires a thresholding function $\tilde{\psi}$ since the set of allowable actions $K_\mathcal{D}$ is given, hence no action need to be explicitly truncated.
% Specifically, we propose to modify the learned policy form as 
\begin{align}
    \pi_{t+1, \datasetPolicy}(a|s) = \exp_q\AdaBracket{\frac{Q_{t, \datasetPolicy}(s,a)}{\tau} - \tilde{\psi}{\AdaBracket{\frac{Q_{t, \datasetPolicy}(s,\cdot)}{\tau}}}},  \quad \sum_{a\in K_{\mathcal{D}}(s)} \pi_{t+1, \datasetPolicy} = 1.
    \label{eq:proposal_policy}
\end{align}
The thresholding function together with the given $K_\mathcal{D}$ implies during the backward learning the new set of allowable actions $K_{t, q}$, where we make clear the dependence on $q$ and iteration $t$.
The set should satisfy the condition $K_{t, q} \preceq K_\mathcal{D}$:
 i.e. the cardinality $|K_{t, q}| \leq |K_\mathcal{D}|$ and support constraint $\pi_t \preceq \datasetPolicy$.
Another advantage lies in that for $q>1$, the policy is a variant of categorical distributions which is less susceptible to numerical issues than exponential functions \cite{Tsai2021-selfsupervisedRelativePredictiveCoding}.

@@

\textbf{Why use Tsallis backward learning? }
If the Tsallis behavior policy assumption holds, and Tsallis backward learning is adopted, then the KL divergence between the learned policy and the behavior policy can be uppper-bounded.
That being said, the learned policy is guaranteed to better respect the behavior policy.
\begin{theorem}
    Suppose the dataset $\mathcal{D}$ is generated by a Tsallis behavior policy of any entropic index $q^*$ (it is not important).
    Let $K_{t, q}(s) \preceq K_\mathcal{D}(s)$ denote the set of allowable actions at $t$-th iteration whose cardinality is smaller than $K_\mathcal{D}(s)$.
    Also let $\pi_{t}(a|s) \propto \exp_q\AdaBracket{\frac{Q_{t-1}(s,a)}{\tau}}$ denote the learned policy. 
    Then the KL divergence between the $\pi_t$ and the behavior policy can be upper bounded:
\begin{align}
    \begin{split}
    &\KLany{\pi_t(\cdot|s)}{\datasetPolicy(\cdot|s)}  \leq  |K_{t}(s)|\! \AdaRectBracket{  { \pi_{t}^{q}(a|s) - \pi_{t}^{q-1} (a|s) } + \pi_t(a|s) - \frac{q-3}{q-1}  + \datasetPolicy^{q-2}(a|s) - \datasetPolicy^{q-1}(a|s)}.
\end{split}
\end{align}
\end{theorem}
The bound is suggesting.
It explains why using in-sample softmax $q=1$ may not be a good choice as in this case the KL divergence to the behavior policy is unbounded.
On the other hand, choosing $q=2$ which recovers sparsemax is bounded by $K_{t, q}(s)\AdaBracket{\pi_t(a|s) - \pi_\mathcal{D}(a|s) + 2}$.
However, it should be noted that there is a trade-off between the power of policies $\pi^{q}$ and the cardinality of $K_{t,q}(s)$: $K_{t,q}(s)$ tends to collect all actions when $q \rightarrow \infty$.
\begin{proof}
    Before we state the main result, we prove the following lemma:
\begin{lemma}\label{lemma:log_qlog_diff}
    The difference between the standard logarithm and $q$-logarithm can be expressed by:
    \begin{align*}
        \ln x - \ln_q x = (q-1) \AdaRectBracket{\frac{d}{dq}\ln_q x - \ln x \ln_q x}.
    \end{align*}
\end{lemma}
\begin{proof}
    Let us begin with the right hand side 
    \begin{align*}
        &(q-1) \AdaRectBracket{\frac{d}{dq}\ln_q x - \ln x \ln_q x} = (q-1) \AdaRectBracket{\frac{d}{dq} \frac{x^{q-1} - 1}{q-1} -  \ln x \ln_q x}\\
        &= (q-1) \AdaRectBracket{ \frac{(x^{q-1} - 1)'(q-1) - (x^{q-1} - 1)(q-1)'}{(q-1)^2  } -  \ln x \ln_q x }\\
        &= (q-1) \AdaRectBracket{ \frac{(q-1) x^{q-1}\ln x - (x^{q-1} - 1)}{(q-1)^2  } -  \ln x \ln_q x }\\
        &= x^{q-1}\ln x - \ln_q x - (q-1)\ln x\ln_q x = \AdaBracket{(q-1)\ln_q x + 1} \ln x - \ln_q x - (q-1)\ln x\ln_q x\\
        &= \ln x - \ln_q x.
    \end{align*}
\end{proof}
With the lemma, we have the following theorem indicating the Tsallis backward learning  policies have bounded distance to the behavior policy:
    We decompose the KL divergence into three terms and apply Lemma \ref{lemma:log_qlog_diff}:
\begin{align}
    \begin{split}
        &\KLany{\pi_t(\cdot|s)}{\datasetPolicy(\cdot|s)} = \expectation{a\sim\pi_{t}(\cdot | s)}{\ln \pi_{t}(a|s) - \ln\datasetPolicy(a|s)} \\
        & = \expectation{a\sim\pi_{t}(\cdot | s)}{\underbrace{\ln \pi_{t}(a|s) - \ln_q\pi_{t}(a|s)}_{(1)} + \underbrace{\ln_q\pi_{t}(a|s) -  \ln_q\datasetPolicy(a|s)}_{(2)} + \underbrace{\ln_q\datasetPolicy(a|s) - \ln\datasetPolicy(a|s)}_{(3)} } .
    \end{split}
\end{align}
Let us now respectively bound the three terms:
\begin{align}
    \begin{split}
        (1): &\ln \pi_{t}(a|s) - \ln_q\pi_{t}(a|s) = (q-1) \AdaRectBracket{\frac{d}{dq}\ln_q \pi_t(a|s) - \ln_q\pi_t(a|s)\ln\pi_t(a|s)}\\
        % &= (q-1)\AdaRectBracket{\pi_t^{q-1}(a|s)\ln\pi_t(a|s) - \frac{\pi_{t}^{q-1}(a|s)-1}{(q-1)^2} - \ln_q\pi_t(a|s)\ln\pi_t(a|s)}\\
        &= \pi_{t}^{q-1}(a|s)\ln\pi_t(a|s) - \ln_q\pi_t(a|s) - (q-1)\ln_q\pi_t(a|s)\ln\pi_t(a|s)\\
        % &= (q-1)\pi_{t}^{q-1}(a|s)\ln\pi_t(a|s) - \AdaBracket{ \frac{Q_{t-1}(s,a)}{\tau} - \tilde{\psi}\AdaBracket{\frac{Q_{t-1}(s,\cdot)}{\tau}}}  + \ln\pi_t(a|s) \AdaBracket{ \frac{Q_{t-1}(s,a)}{\tau} - \tilde{\psi}\AdaBracket{\frac{Q_{t-1}(s,\cdot)}{\tau}}}
        & \leq \pi_{t}^{q-1}(a|s)\ln\pi_t(a|s) + \frac{1}{q-1} + \ln\pi_t (a|s)\\
        & \leq  \AdaBracket{ \pi_{t}^{q}(a|s) - \pi_{t}^{q-1} (a|s) } + \pi_t(a|s) - \frac{q-2}{q-1},
    \end{split}
\end{align}
where we leveraged $\ln x \leq x - 1$ and  the fact that $\forall a\in K(s), \,\, -\frac{1}{q-1} \leq \frac{Q_{t-1}(s,a)}{\tau} - \tilde{\psi}\AdaBracket{\frac{Q_{t-1}(s,\cdot)}{\tau}} \leq 0$. 
If $a \notin K(s)$, then $\ln\pi_t(a|s) = -\infty$ and the KL term is unbounded.
The same fact is exploited to yield an upper bound $\frac{1}{q-1}$  for (2).
We now work with (3):
\begin{align}
    \begin{split}
        (3): &\ln_q \datasetPolicy(a|s) - \ln\datasetPolicy(a|s) = -(q-1) \AdaRectBracket{\frac{d}{dq}\ln_q \datasetPolicy(a|s) - \ln_q\datasetPolicy(a|s)\ln\datasetPolicy(a|s)}\\
        % &= -\pi_{t}^{q-1}(a|s)\ln\pi_t(a|s) + \ln_q\pi_t(a|s) + (q-1)\ln_q\pi_t(a|s)\ln\pi_t(a|s)\\
        & \leq -\datasetPolicy^{q-1}(a|s)\ln\datasetPolicy(a|s)  \leq  -\datasetPolicy^{q-1}(a|s)\AdaBracket{1 - \frac{1}{\datasetPolicy(a|s)} } = \datasetPolicy^{q-2}(a|s) - \datasetPolicy^{q-1}(a|s).
    \end{split}
\end{align}
Putting all terms together, we arrive at the upper bound that 
\begin{align*}
    \KLany{\pi_t(\cdot|s)}{\datasetPolicy(\cdot|s)} \leq K_{t}(s)\! \AdaRectBracket{  { \pi_{t}^{q}(a|s) - \pi_{t}^{q-1} (a|s) } + \pi_t(a|s) - \frac{q-3}{q-1}  + \datasetPolicy^{q-2}(a|s) - \datasetPolicy^{q-1}(a|s)}.
\end{align*}

\end{proof}

@@

Furthermore, the assumption naturally leads to the choice of Tsallis regularized backward learning which recovers in-sample softmax as a special case when $q=1$.
Since softmax policies have full support, there always exists non-negligible probabilities for suboptimal actions; while for \eq{\ref{eq:proposal_policy}} the optimal policy can switch between greedy policy and multimodal policy.
The former is achieved when all but one action have values lower than the threshold; while vice versa for the latter.

@@

We follow the same setup here, but replacing every appearance of $\ln, \exp$ to their \qlog and $q$-exponential counterpart.
In practice, computing the thresholding function $\psi$ requires sorting in the discrete control setting and it remains unclear how to accurately estimate it for the continuous control problems.
We found it improves the performance by simply removing $\psi$ and renormalize the policy.
In this case, thresholding can still be achieved by simply learning $Q_{\theta} \rightarrow 0$.
That is, the learned policy form becomes $\pi_{t+1}(a|s) \propto \AdaRectBracket{Q_{\theta}(s,a)}_{+}$.
In order to also sample actions from the dataset when updating the policy, we repeat the derivation in \eq{\ref{eq:insample_policy}} but for the sparsemax policy.
% Let us define $q_{\theta}(s,a) := \frac{Q_{\datasetPolicy, k}(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_{\datasetPolicy, k}(s,\cdot)}{\tau}}$ for convenience:
\begin{align}
    \begin{split}
        \pi_{\datasetPolicy, k+1} (a|s) &= \datasetPolicy(a|s) \datasetPolicy(a|s)^{-1} \exp_q{Q_{\theta}(s,a)} \\
        &= \datasetPolicy(a|s)  \exp_q{\AdaBracket{\ln_q{\frac{1}{\datasetPolicy(a|s)}}}} \exp_q{Q_{\theta}(s,a)}\\
        &=  \datasetPolicy(a|s) \sqrt[q-1]{\AdaBracket{ \exp_q\AdaBracket{Q_{\theta}(s,a) + \ln_2\frac{1}{\datasetPolicy(a|s)}}^{q-1} - (q-1)^2 Q_{\theta}(s,a)  \ln_q{\frac{1}{\datasetPolicy(a|s)}} }}.
        \label{eq:tsallis_inac_policy}
    \end{split}
\end{align}
In the last step we made use of the relationship $\AdaBracket{\expq{x}\cdot \expq{y}}^{q-1} = \expq{\AdaBracket{x+y}}^{q-1} + (q-1)^2 xy$ \cite{Yamano2004-properties-qlogexp}.
Similar to the discussion after \eq{\ref{eq:inac_policy}}, the Tsallis policy we derived here can be seen as the result from Tsallis KL regularization, plus another regularization that gives rise to the additional term inside the $\exp_q$ function.
In the implementation, we choose the sparsemax parametrization, which gives the following Tsallis in-sample sparsemax actor-critic update rule:
% In the discrete case, the truncation threshold or the normalization function $\psi$ is computed by sorting the action values.
% However, in the continuous case we can no longer do that.
% Instead, we propose to use a single-point estimate:
% \begin{align*}
%     &V(s) \approx \frac{1}{2}\tau \AdaBracket{ \AdaBracket{\frac{Q_{\datasetPolicy, k}(s',a')}{\tau}}^2 - \psi \AdaBracket{\frac{Q_{\datasetPolicy, k}(s', \cdot)}{\tau}}^2 + 1} \\
%     &\Leftrightarrow \psi\AdaBracket{\frac{Q_{\datasetPolicy, k}(s,\cdot)}{\tau}} \approx \sqrt{\AdaBracket{\frac{Q_{\datasetPolicy, k}(s',a')}{\tau}}^2 - \frac{V(s) - \frac{1}{2}\tau}{\frac{1}{2}\tau}}.
% \end{align*}
% {\color{red}(need to record this $\psi$ value in the experiment to see how big area of the action Gaussian distribution is truncated.)}
% The single-point estimate is similar to the common practice of entropy-regularized actor-critic algorithms such as soft actor-critic \cite{haarnoja-SAC2018} that approximates the value function by a single-point estimate, i.e. $V_{\zeta}(s) \approx Q_{\theta}(s,a) + \ln\pi_{\phi}(a|s)$.
\begin{align}
    % & \psi\AdaBracket{\frac{Q_{\datasetPolicy, \theta}(s,\cdot)}{\tau}} \leftarrow \sqrt{\AdaBracket{\frac{Q_{\theta}(s,a)}{\tau}}^2 - \frac{V_{\zeta}(s) - \frac{\tau}{2}}{\frac{\tau}{2}} }, \\ 
    % & Q_{\theta}(s,a) \leftarrow \frac{Q_{\theta}(s,a)}{\tau} - \psi\AdaBracket{\frac{Q_{\theta}(s,\cdot)}{\tau}},\\
    &\mathcal{L}_\text{actor}(\phi) = -\expectation{s,a \sim \mathcal{D}}{ \AdaBracket{ \exp_q\AdaBracket{Q_{\theta}(s,a) + \ln_2\frac{1}{\pi_{\omega}(a|s)}} -  Q_{\theta}(s,a)  \logq{\frac{1}{\pi_{\omega}(a|s)}} } \ln\pi_{\phi}(a|s) }, \\
    & \mathcal{L}_\text{baseline}(\zeta) = \expectation{s\sim\mathcal{D}, a\sim \pi_{\phi}(s)}{\AdaBracket{ v_{\zeta}(s) - Q_{\theta}(s,a) - \tau\ln_2\pi_{\phi}(a|s)}^2},\\
    & \mathcal{L}_\text{critic}(\theta) = \expectation{s,a,r,s' \sim \mathcal{D}}{\AdaBracket{r + \gamma v_{\zeta}(s') - Q_{\theta}(s,a)}^2 }.
\end{align}
